\documentclass{article}

% Encodings, page setup, paragraph formatting, font
\usepackage[top=0.9in, bottom=1in, left=1.5in, right=1.5in]{geometry}
\usepackage[icelandic]{babel}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage[parfill]{parskip}
\usepackage{cancel}
\usepackage{comment}
% Tables and lists
\usepackage{booktabs,tabularx}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage{adjustbox}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{xcolor}
% Math
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{gensymb}
% Graphics
\usepackage{graphicx}
\usepackage{forest}
\usepackage{tikz}
\usetikzlibrary{positioning, shapes, arrows.meta}
% Code environment
\usepackage{listingsutf8}
\definecolor{commentcolor}{RGB}{0, 128, 0} % Grænn
\definecolor{keywordcolor}{RGB}{0, 0, 255}   % Blár
\definecolor{stringcolor}{RGB}{163, 21, 21}  % Dökkrauður
\definecolor{numbercolor}{RGB}{128, 0, 128}  % Fjólublár
\definecolor{identifiercolor}{RGB}{0, 0, 0}  % Svartur

\def\ojoin{\setbox0=\hbox{$\bowtie$}%
  \rule[-.02ex]{.25em}{.4pt}\llap{\rule[\ht0]{.25em}{.4pt}}}
\def\leftouterjoin{\mathbin{\ojoin\mkern-5.8mu\bowtie}}
\def\rightouterjoin{\mathbin{\bowtie\mkern-5.8mu\ojoin}}
\def\fullouterjoin{\mathbin{\ojoin\mkern-5.8mu\bowtie\mkern-5.8mu\ojoin}}

\lstset{
    language=Java,
    basicstyle=\ttfamily,
    keywordstyle=\color{keywordcolor}\bfseries,
    commentstyle=\color{commentcolor},
    identifierstyle=\color{identifiercolor},
    stringstyle=\color{stringcolor},
    showstringspaces=false,
    numbers=left,
    numberstyle=\tiny\color{gray},
    tabsize=2,
    breaklines=true,
    columns=fullflexible,
    keepspaces=true,
    inputencoding=utf8, 
    extendedchars=true,
    literate=
        {á}{{\'a}}1
        {ð}{{\dh}}1
        {é}{{\'e}}1
        {í}{{\'i}}1
        {ó}{{\'o}}1
        {ú}{{\'u}}1
        {ý}{{\'y}}1
        {þ}{{\th}}1
        {æ}{{\ae}}1
        {ö}{{\"o}}1
        {Á}{{\'A}}1
        {Ð}{{\DH}}1
        {É}{{\'E}}1
        {Í}{{\'I}}1
        {Ó}{{\'O}}1
        {Ú}{{\'U}}1
        {Ý}{{\'Y}}1
        {Þ}{{\TH}}1
        {Æ}{{\AE}}1
        {Ö}{{\"O}}1,
}

% Restin af forskriftinni
\usepackage[pdftex,bookmarks=true,colorlinks=true,pdfauthor={Hafsteinn Einarsson},
            linkcolor=blue,urlcolor=blue]{hyperref}

%Custom Commands til að auðvelda mér lífið
\newcommand{\sv}{\textbf{Svar: }}
\newcommand{\bo}[1]{\textbf{#1}}
\newcommand{\enum}{\begin{enumerate}[label = \alph*.]}

% Hyphenation
\hyphenpenalty=5000
% Page and section numbering
\setcounter{secnumdepth}{-1} 
\pagenumbering{gobble}

\title{Verkefni 20}
\author{brj46 }
\date{vor 2025}

\begin{document}

\maketitle

\bo{Consider a system with a Paged Memory Management Unit (PMMU) that supports 12 bit
physical addresses, however a process is only able to use 10 bit logical addresses. The page
size is 64 byte, the word size is 1 byte.}

\begin{enumerate}
    \item Into how many pages is the logical address space divided?\\
    \sv{$2^{10} \div 2^6 = 2^4 = 16$ pages}

    \item Into how many frames is the physical address space divided?\\
    \sv{$2^{12} \div 2^6 = 2^6 = 64$ frames}

    \item What is the maximum degree of multiprogramming (how many programs can be loaded) if no swapping/demand paging is used?\\
    \sv{Each process could use all 16 of its pages. There are 64 frames, so 4 processes could be loaded at once.}

    \item Is it possible to state how many processes can be in the system simultaneously if swapping/demand paging is used? Explain.\\
    \sv{No. With swapping/demand paging, not all pages need to reside in physical memory at once, so more than 4 processes can be loaded.}

    \item How many entries does a page table have in this system?\\
    \sv{Each page table has 16 entries (one per page).}

    \item How many bits are needed for each page table entry (assuming only the minimal number for the frame number plus valid bit)?\\
    \sv{We need 6 bits for the frame number (since there are 64 frames) and 1 bit for the valid bit, so 7 bits in total.}

    \item Given the first four frame numbers in the page table are 7, 1, 10, 12 (the first entry = page 0, second entry = page 1, etc.), what is the physical address for logical address 42?\\
    \sv{The page number is 0 (since 42 < 64), and offset is 42. Page 0 maps to frame 7, so the physical address is $7 \times 64 + 42 = 490$.}

    \item What logical address maps to physical address 681?\\
    \sv{$681 \div 64 = 10$ (integer division), remainder $41$, so the frame is 10, the offset is 41. 
    From the table, frame 10 corresponds to page 2. Hence the logical address is $2 \times 64 + 41 = 169$.}

\end{enumerate}

\end{document}
